GG_topi

### 2018-11-19 gg_topi.Rmd
T7 no 24/7

---
title: "GG_topi2"
author: "Francesco Beghini"
date: "26/09/2018"
output: 
  html_document: 
    theme: cerulean
editor_options: 
  chunk_output_type: console
---

@@

```{r load qiime, include=FALSE, cache=TRUE}
gg_topi <- import_biom(BIOMfilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus_silva/otu_table_mc10_w_tax.biom',
                         treefilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus_silva/rep_set.tre',
                         refseqfilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus_silva/rep_set.fna',
                        refseqFunction = parse_taxonomy_default
                      )

@@

@@

# Beta diversity
```{r Beta diversity, cache=TRUE}
distwu <- phyloseq::distance(gg_topi, "wunifrac")
distbc <- phyloseq::distance(gg_topi, "bray")
distjac <- phyloseq::distance(gg_topi, "jaccard")
distwu_meta <- phyloseq::distance(gg_topi_meta_samples, "jaccard")

@@

## Beta diversity between timepoints, T1 is the baseline
```{r Beta diversity between timepoints t1}
distances[distances$timepoint1 == 1 | distances$timepoint1 == 2, 'timepoint1'] <- 2
distances[distances$timepoint2 == 1 | distances$timepoint2 == 2, 'timepoint2'] <- 2

t1 <- distances %>% 
  filter(topo1==topo2) %>%
  filter(timepoint1==2) %>%
  filter(timepoint1!=timepoint2) %>%
  arrange(topo1)

@@

## Differential analysis
```{r DA, cache=TRUE}
threshold <- 0.05

ggtopi_filt <- filter_taxa(gg_topi, function(x) mean(x) > 5e-1, TRUE)

dds <- DESeq2::DESeq(phyloseq_to_deseq2(ggtopi_filt, design = ~ cond_vac), parallel = TRUE, quiet = TRUE, fitType = 'parametric')

mmtx <- model.matrix(object = ~timepoint*real_vaccine*real_bifido, data = data.frame(gg_topi@sam_data))
idx <- which(apply(mmtx, 2, function(x) all(x==0)))
mmtx <- mmtx[,-idx]

rmmtx <- model.matrix(object = ~ timepoint, data = data.frame(gg_topi@sam_data))

pds <- phyloseq_to_deseq2(ggtopi_filt, design = ~timepoint*real_vaccine*real_bifido, ignoreRank = TRUE)

dds <- DESeq2::DESeq(pds, full = mmtx, reduced = rmmtx, test = 'LRT')

dds_bif <- DESeq2::DESeq(phyloseq_to_deseq2(gg_topi, design = ~ cond_bif), parallel = TRUE, quiet = TRUE)
dds1 <- DESeq2::DESeq(phyloseq_to_deseq2(ggtopi_filt, design = ~ timepoint), parallel = TRUE, quiet = TRUE)

dds_phylum <- DESeq2::DESeq(phyloseq_to_deseq2(gg_topi.phylum, design = ~ timepoint), parallel = TRUE, quiet = TRUE)

```
## T2 vs T3
```{r DA T2 vs T3}
res23 <- DESeq2::results(dds1, contrast = c("timepoint","T2","T3"), alpha = 0.05, format = "DataFrame")
res23.filtered <- res23[!is.na(res23$padj),]
res23.filtered <- res23.filtered[res23.filtered$padj < threshold,]
res23.filtered <- res23.filtered[abs(res23.filtered$log2FoldChange) > 2,]

res23.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res23.filtered),])) %>%
  ggplot(aes(log2FoldChange, Genus)) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = log(baseMean))) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  ggtitle('T1 vs T2') +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

## T2 vs T3 phylum level
```{r DA T2 vs T3 phylum}
res_phylum <- DESeq2::results(dds_phylum, contrast = c("timepoint","T2","T3"), alpha = 0.05, format = "DataFrame")
res_phylum.filtered <- res_phylum[!is.na(res_phylum$padj),]
res_phylum.filtered <- res_phylum.filtered[res_phylum.filtered$padj < threshold,]
# res_phylum.filtered <- res_phylum.filtered[abs(res_phylum.filtered$log2FoldChange) > 2,]

res_phylum.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res_phylum.filtered),])) %>%
  ggplot(aes(log2FoldChange, Phylum)) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  theme_bw() +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

## Check T1 vs T2
```{r DA T1 vs T2}
res12 <- DESeq2::results(dds1, contrast = c("timepoint","T1","T2"))
res12.filtered <- res12[!is.na(res12$padj),]
res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]

res12.filtered %>% 
  as.data.frame %>% 
  cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
  ggplot(aes(log2FoldChange, Genus)) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  ggtitle('Timepoint 1 vs Timepoint 2') +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

```{r DA T3 vs T4}
res23 <- DESeq2::results(dds1, contrast = c("timepoint","T3","T4"))
res23.filtered <- res23[!is.na(res23$padj),]
res23.filtered <- res23.filtered[res23.filtered$padj < threshold,]

res23.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res23.filtered),])) %>%
  ggplot(aes(log2FoldChange, Genus)) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  ggtitle('Timepoint 3 vs Timepoint 4') +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
)
```

## Differential analysis Pentatome T7 vs not vaccinated T7
```{r}
  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T7_YES","T7_NO"))
  res12 <- DESeq2::results(dds)
  res12.filtered <- res12[!is.na(res12$padj),]
  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]

  plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
    ggplot(aes(log2FoldChange, Genus)) +
    geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
    geom_point(size=3, aes(color = log10(baseMean))) +
    facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
    theme_bw() +
    ggtitle('Pentatome T3 vs T7') +
    scale_y_discrete(position = "right") +
    theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
          strip.text = element_text(size = 9),
          axis.title = element_text(size = 9),
          axis.title.y = element_blank(),
          strip.background = element_rect(colour = "transparent", fill = "transparent"),
          strip.text.y = element_text(angle = 180)
    ))

## Differential analysis Pentatome T3 VS T7
```{r}
  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T7_YES","T3_NO"))
  res12.filtered <- res12[!is.na(res12$padj),]
  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]

  plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
    ggplot(aes(log2FoldChange, Genus)) +
    geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
    geom_point(size=3, aes(color = log10(baseMean))) +
    facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
    theme_bw() +
    ggtitle('Pentatome T3 vs T7') +
    scale_y_discrete(position = "right") +
    theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
          strip.text = element_text(size = 9),
          axis.title = element_text(size = 9),
          axis.title.y = element_blank(),
          strip.background = element_rect(colour = "transparent", fill = "transparent"),
          strip.text.y = element_text(angle = 180)
    ))

```{r}
res12.filtered %>% 
  as.data.frame %>% 
  cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
  filter(is.na(Phylum)) %>%  knitr::kable()
```

## Differential analysis Empty vescicles T3 VS T7
```{r}
  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T7_NO","T3_NO"))
  res12.filtered <- res12[!is.na(res12$padj),]
  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]

  plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
    ggplot(aes(log2FoldChange, Genus)) +
    geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
    geom_point(size=3, aes(color = baseMean)) +
    facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
    theme_bw() +
    ggtitle('Empty T3vsT7') +
    scale_y_discrete(position = "right") +
    theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
          strip.text = element_text(size = 9),
          axis.title = element_text(size = 9),
          axis.title.y = element_blank(),
          strip.background = element_rect(colour = "transparent", fill = "transparent"),
          strip.text.y = element_text(angle = 180)
    ))

## Differential analysis Pentatoma vs empty vescicles T7
```{r}
  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T7_YES","T7_NO"))
  res12.filtered <- res12[!is.na(res12$padj),]
  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]

  plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
    ggplot(aes(log2FoldChange, Genus)) +
    geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
    geom_point(size=3, aes(color = baseMean)) +
    facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
    theme_bw() +
    ggtitle(paste('PBS vs Pentatoma at timepoint', timepoint_number, sep = ' ')) +
    scale_y_discrete(position = "right") +
    theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
          strip.text = element_text(size = 9),
          axis.title = element_text(size = 9),
          axis.title.y = element_blank(),
          strip.background = element_rect(colour = "transparent", fill = "transparent"),
          strip.text.y = element_text(angle = 180)
    ))

## Differential analysis Pentatoma vs empty vescicles T3
```{r}
  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T3_YES","T3_NO"))
  res12.filtered <- res12[!is.na(res12$padj),]
  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]
  timepoint_number <- stringr::str_sub(timepoint,17,17)

  plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
    ggplot(aes(log2FoldChange, Genus)) +
    geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
    geom_point(size=3, aes(color = baseMean)) +
    facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
    theme_bw() +
    ggtitle(paste('PBS vs Pentatoma at timepoint', timepoint_number, sep = ' ')) +
    scale_y_discrete(position = "right") +
    theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
          strip.text = element_text(size = 9),
          axis.title = element_text(size = 9),
          axis.title.y = element_blank(),
          strip.background = element_rect(colour = "transparent", fill = "transparent"),
          strip.text.y = element_text(angle = 180)
    ))

## Differential analysis Pentatome vs No-pentatome at different timepoints
```{r}
for (timepoint in c("merge_timepointT3.pentatomePBS","merge_timepointT4.pentatomePBS","merge_timepointT5.pentatomePBS","merge_timepointT6.pentatomePBS","merge_timepointT7.pentatomePBS")) {
  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T6_Pentatoma","T6_Empty"), )
  res12.filtered <- res12[!is.na(res12$padj),]
  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]
  timepoint_number <- stringr::str_sub(timepoint,17,17)

  plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
    ggplot(aes(log2FoldChange, Genus)) +
    geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
    geom_point(size=3, aes(color = baseMean)) +
    facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
    theme_bw() +
    ggtitle(paste('PBS vs Pentatoma at timepoint', timepoint_number, sep = ' ')) +
    scale_y_discrete(position = "right") +
    theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
          strip.text = element_text(size = 9),
          axis.title = element_text(size = 9),
          axis.title.y = element_blank(),
          strip.background = element_rect(colour = "transparent", fill = "transparent"),
          strip.text.y = element_text(angle = 180)
    ))
}
```

<!-- ## Differential analysis Bifido vs No-bifido at different timepoints -->
<!-- ```{r} -->
<!-- for (timepoint in c("bifidoPBS.merge_timepointT3","bifidoPBS.merge_timepointT4","bifidoPBS.merge_timepointT5","bifidoPBS.merge_timepointT6","bifidoPBS.merge_timepointT7")) { -->
<!--   res12 <- DESeq2::results(dds2, contrast = list(c("bifido_PBS_vs_Bifido" , timepoint))) -->
<!--   res12.filtered <- res12[!is.na(res12$padj),] -->
<!--   res12.filtered <- res12.filtered[res12.filtered$padj < threshold,] -->
<!--   timepoint_number <- stringr::str_sub(timepoint,-1) -->

<!--   plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>% -->
<!--     ggplot(aes(log2FoldChange, Genus)) + -->
<!--     geom_vline(xintercept = 0.0, color = "gray", size = 0.5) + -->
<!--     geom_point(size=3, aes(color = baseMean)) + -->
<!--     facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+ -->
<!--     theme_bw() + -->
<!--     ggtitle(paste('Bifido vs no bifido at timepoint', timepoint_number, sep = ' ')) + -->
<!--     scale_y_discrete(position = "right") + -->
<!--     theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11), -->
<!--           strip.text = element_text(size = 9), -->
<!--           axis.title = element_text(size = 9), -->
<!--           axis.title.y = element_blank(), -->
<!--           strip.background = element_rect(colour = "transparent", fill = "transparent"), -->
<!--           strip.text.y = element_text(angle = 180) -->
<!--     )) -->
<!-- } -->
<!-- ``` -->

# PERMANOVA
```{r}
metadata <- as(sample_data(gg_topi), "data.frame")
permanova.res <- data.frame()
distwu <- phyloseq::distance(gg_topi, "wunifrac")

@@

```{r}
for(s in combn(levels(metadata$merge_timepoint),2, simplify = FALSE)){
  metadata_2cat <- metadata[metadata$merge_timepoint %in% s,]
  distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, merge_timepoint %in% s), "wunifrac")
  permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)
  permanova.res %<>% bind_rows(data.frame(contrast = stringr::str_c(s, collapse = " vs "), r2=permanova$aov.tab[1,5],pvalue=permanova$aov.tab[1,6])) %>% arrange(pvalue)
}
```

```{r}
for(s in combn(levels(metadata$pentatome),2, simplify = FALSE)){
  metadata_2cat <- metadata[metadata$pentatome %in% s,]
  distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, pentatome %in% s), "wunifrac")
  permanova <- vegan::adonis(formula = distwu_2cat ~ pentatome, data = metadata_2cat)
  permanova.res %<>% bind_rows(data.frame(contrast = stringr::str_c(s, collapse = " vs "), r2=permanova$aov.tab[1,5],pvalue=permanova$aov.tab[1,6])) %>% arrange(pvalue)
}
```

```{r}
for(s in combn(levels(metadata$bifido),2, simplify = FALSE)){
  metadata_2cat <- metadata[metadata$bifido %in% s,]
  distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, bifido %in% s), "wunifrac")
  permanova <- vegan::adonis(formula = distwu_2cat ~ bifido, data = metadata_2cat)
  permanova.res %<>% bind_rows(data.frame(contrast = stringr::str_c(s, collapse = " vs "), r2=permanova$aov.tab[1,5],pvalue=permanova$aov.tab[1,6])) %>% arrange(pvalue)
}
```

@@

## PERMANOVA all vs all
```{r}
permanova.res <- data.frame()
for (s in combn(levels(sample_data(gg_topi)$cond_all),2, simplify = FALSE)) {
  metadata_2cat <- data.frame(sample_data(gg_topi))[sample_data(gg_topi)$cond_all %in% s,]
  distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, cond_all %in% s), "wunifrac")
  permanova <- vegan::adonis(formula = distwu_2cat ~ cond_all, data = metadata_2cat)
  permanova.res %<>% bind_rows(data.frame(contrast = stringr::str_c(s, collapse = " vs "), r2=permanova$aov.tab[1,5],pvalue=permanova$aov.tab[1,6])) %>% arrange(pvalue)
}
```

@@

## Correlation beta-diversity T2-TX, tumor mass
```{r corr tumor wunifrac}
distances %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==2) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description')) %>%
  ggplot(aes(value,tumor_mass)) +
  geom_point() + 
  facet_grid(timepoint2~condizione, scales = "free", space = "free_y")+
  geom_smooth(method = 'lm')
```

```{r}
wunif_ds <- distances %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==2) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description'))

@@

```{r corr tumor braycurtis}
distancesbc %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==2) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description')) %>%
  ggplot(aes(value,tumor_mass)) +
  geom_point() + 
  stat_smooth_func(geom="text",method="lm",hjust=0,parse=TRUE) +
  facet_grid(timepoint2~condizione, scales = "free", space = "free_y")+
  geom_smooth(method = 'lm')
```

```{r}
brayc_ds <- distancesbc %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==2) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description'))

@@

```{r corr tumor jaccard}
distancesjac %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==2) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description')) %>%
  ggplot(aes(value,tumor_mass)) +
  geom_point() + 
  stat_smooth_func(geom="text",method="lm",hjust=0,parse=TRUE) +
  facet_grid(timepoint2~condizione, scales = "free", space = "free_y")+
  geom_smooth(method = 'lm')
```

```{r}
jac_ds <- distancesjac %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==2) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description'))

@@

# Tracking OTU
```{r}
counts <- data.frame(as.data.frame(otu_table(gg_topi)[c('EU622777.1.1487','EU504401.1.1405'),] %>% t)) %>% tibble::rownames_to_column('sampleid')
colnames(counts) <- c('sampleid',"Lachnospiraceae_NK4A136","Muribaculaceae")
counts$t <- sapply(as.character(counts$sampleid), function(x) strsplit(x,'-')[[1]][4])
counts$condizione <- sapply(as.character(counts$sampleid), function(x) strsplit(x,'-')[[1]][2])

reshape2::melt(counts, id.vars = c('t', 'condizione','sampleid')) %>%
  ggplot() +
  stat_boxplot(aes(t, value)) + 
  facet_grid(variable ~ condizione, scales = "free_x", space="free_y")
```

>>>>

---
title: "GG_topi"
author: "Francesco Beghini"
date: "26/09/2018"
output: 
  html_document: 
    df_print: kable
    self_contained: no
    theme: cerulean
editor_options: 
  chunk_output_type: inline
---

@@

```{r load qiime, include=FALSE, cache=TRUE}
gg_topi <- import_biom(BIOMfilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus/otu_table_mc10_w_tax_rdp.biom',
                         treefilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus/rep_set.tre',
                         refseqfilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus/rep_set.fna',
                        refseqFunction = parse_taxonomy_default
                      )

@@

unass_otus <- tax_table(gg_topi) %>%
                data.frame %>%
                tibble::rownames_to_column('OTU') %>%
                filter(Domain == 'Unassigned') %>%
                select(OTU)

unknown_otus <- tax_table(gg_topi) %>%
                data.frame %>%
                tibble::rownames_to_column('OTU') %>%
                filter(is.na(Genus)) %>%
                select(OTU)

tax_table(gg_topi)[unass_otus$OTU, ] <- rep.int('Unassigned', 7)
tax_table(gg_topi)[unknown_otus$OTU, 'Genus'] <- 'Unknown'

@@

# Beta diversity
```{r Beta diversity, cache=FALSE}
distwu <- phyloseq::distance(gg_topi, "wunifrac")
distbc <- phyloseq::distance(gg_topi, "bray")
distjac <- phyloseq::distance(gg_topi, "jaccard")
distwu_meta <- phyloseq::distance(gg_topi_meta_samples, "jaccard")

@@

## Beta diversity between timepoints, T1 is the baseline
```{r Beta diversity between timepoints t1}
distances_tmp <- distances
distances_tmp[distances_tmp$timepoint1 == 1 | distances_tmp$timepoint1 == 2, 'timepoint1'] <- 2
distances_tmp[distances_tmp$timepoint2 == 1 | distances_tmp$timepoint2 == 2, 'timepoint2'] <- 2

t1 <- distances_tmp %>% 
  filter(topo1==topo2) %>%
  filter(timepoint1==2) %>%
  filter(timepoint1!=timepoint2) %>%
  arrange(topo1)

@@

# PERMANOVA
```{r, cache=TRUE}
metadata <- as(sample_data(gg_topi), "data.frame")
permanova.res <- data.frame()
distwu <- phyloseq::distance(gg_topi, "wunifrac")

@@

```{r, cache=TRUE}
for(s in combn(levels(metadata$merge_timepoint),2, simplify = FALSE)){
  metadata_2cat <- metadata[metadata$merge_timepoint %in% s,]
  distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, merge_timepoint %in% s), "wunifrac")
  permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)
  permanova.res %<>% bind_rows(data.frame(contrast = stringr::str_c(s, collapse = " vs "), r2=permanova$aov.tab[1,5],pvalue=permanova$aov.tab[1,6])) %>% arrange(pvalue)
}
```

```{r, cache=TRUE}
for(s in combn(levels(metadata$pentatome),2, simplify = FALSE)){
  metadata_2cat <- metadata[metadata$pentatome %in% s,]
  distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, pentatome %in% s), "wunifrac")
  permanova <- vegan::adonis(formula = distwu_2cat ~ pentatome, data = metadata_2cat)
  permanova.res %<>% bind_rows(data.frame(contrast = stringr::str_c(s, collapse = " vs "), r2=permanova$aov.tab[1,5],pvalue=permanova$aov.tab[1,6])) %>% arrange(pvalue)
}
```

```{r, cache=TRUE}
for(s in combn(levels(metadata$bifido),2, simplify = FALSE)){
  metadata_2cat <- metadata[metadata$bifido %in% s,]
  distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, bifido %in% s), "wunifrac")
  permanova <- vegan::adonis(formula = distwu_2cat ~ bifido, data = metadata_2cat)
  permanova.res %<>% bind_rows(data.frame(contrast = stringr::str_c(s, collapse = " vs "), r2=permanova$aov.tab[1,5],pvalue=permanova$aov.tab[1,6])) %>% arrange(pvalue)
}
```

@@

## PERMANOVA all vs all
```{r, cache=TRUE}
permanova.res <- data.frame()
for (s in combn(levels(sample_data(gg_topi)$cond_all),2, simplify = FALSE)) {
  metadata_2cat <- data.frame(sample_data(gg_topi))[sample_data(gg_topi)$cond_all %in% s,]
  distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, cond_all %in% s), "wunifrac")
  permanova <- vegan::adonis(formula = distwu_2cat ~ cond_all, data = metadata_2cat, parallel = 2)
  permanova.res %<>% bind_rows(data.frame(contrast = stringr::str_c(s, collapse = " vs "), r2=permanova$aov.tab[1,5],pvalue=permanova$aov.tab[1,6])) %>% arrange(pvalue)
}
```

@@

## Correlation beta-diversity T3-TX, tumor mass
```{r corr tumor wunifrac}
distances %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description')) %>%
  ggplot(aes(value,tumor_mass)) +
  geom_point() + 
  facet_grid(timepoint2~condizione, scales = "free", space = "free_y")+
  geom_smooth(method = 'lm')
```

```{r}
wunif_ds <- distances %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description'))

@@

```{r corr tumor braycurtis}
distancesbc %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description')) %>%
  ggplot(aes(value,tumor_mass)) +
  geom_point() +
  facet_grid(timepoint2~condizione, scales = "free", space = "free_y")+
  geom_smooth(method = 'lm')
```

```{r}
brayc_ds <- distancesbc %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description'))

@@

```{r corr tumor jaccard}
distancesjac %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description')) %>%
  ggplot(aes(value,tumor_mass)) +
  geom_point() + 
  facet_grid(timepoint2~condizione, scales = "free", space = "free_y")+
  geom_smooth(method = 'lm')
```

```{r}
jac_ds <- distancesjac %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description'))

@@

```{r, include=FALSE, eval=FALSE}
otus <- data.frame(otu_table(gg_topi.genus.relab), check.names = FALSE) %>% 
  tibble::rownames_to_column('OTU') %>% 
  tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
  tibble::add_column(., timepoint = stringr::str_sub(.$SampleID, 12, 13)) %>%
  tibble::add_column(., mouseID = stringr::str_sub(.$SampleID, 8, 10))

wb<-createWorkbook(type="xlsx")
for(mID in sort(unique(otus$mouseID))){
  tmp <- otus %>%
    filter(mouseID==mID) %>%
    reshape2::dcast(OTU ~ timepoint, value.var = 'Abundance') %>%
    left_join(data.frame(tax_table(gg_topi.genus.relab)[,-7], check.names = FALSE) %>% tibble::rownames_to_column('OTU'))
  sheet <- createSheet(wb, sheetName = mID)
  addDataFrame(tmp, sheet)
}
saveWorkbook(wb, "otutable_topi.xlsx")
```

# Distribution of distances between conditions per timepoint
```{r}
dct <- distances %>%
  filter(timepoint1 == timepoint2) %>%
  select(timepoint1, condizione1, condizione2, value) %>%
  dplyr::rename(timepoint = timepoint1) %>% 
  group_by(timepoint, condizione1, condizione2) %>%
  summarise(mediandistance=median(value)) 
```
```{r include=TRUE,  results = "asis"}
for(t in unique(dct$timepoint)){
  print(dct %>%
    filter(timepoint == t) %>% 
    dplyr::rename(condizione = condizione2) %>%
    spread(condizione1, mediandistance) %>%
    knitr::kable())

}
```

# Distribution of distances between timepoints per conditions
```{r}
dtc <- distances %>%
  filter(condizione1 == condizione2) %>%
  dplyr::rename(condizione = condizione1) %>% 
  group_by(condizione, timepoint1, timepoint2) %>%
  summarise(mediandistance=median(value)) 
```

```{r include=TRUE,  results = "asis"}
for(c in unique(dtc$condizione)){
  print(dtc %>%
    filter(condizione == c) %>% 
    dplyr::rename(timepoint = timepoint1) %>%
    spread(timepoint2, mediandistance) %>%
    knitr::kable())
    cat("\n")
}
```
